---
title: 'Economic Growth and Institutions: Assignment 2'
author: "Mathieu Van der Vel, Abdirahman Nuur and Maximilian Grotz"
date: "1 4 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# Setup

```{r}
library(tidyverse)
```


# Parameters
```{r}
sm <- 0
L_0 <- 7.9
K_0 <- 170
M_0 <- 500
A_0 <- 7
sk <- 0.25
delta <- 0.04
alpha <- 0.33
n <- 0.015
g <- 0.02
x <- 0.7
epsilon <- 0.5
w <- 0.05
k_greek <- 0.75
theta1 <- 0.05
theta2 <- 2
```
# Simulation 1: No abatment effort, `r s_m` = 0
Notes to Mathieu: 
1. I simplifed the paramters seeing the `.` as a decimal point
2. Minor modifications to the formulas like ` y`instead of `Y`, and the some equations
```{r}
sim1 <- data.frame(year = 2020:2100,
                   L = L_0,
                   K = K_0,
                   A = A_0,
                   sm = sm,
                   Z = (K_0^alpha)*((A_0*L_0)^(1-alpha)),
                   m = NA,
                   M = M_0,
                   T = 1 + (M_0^k_greek/1000),
                   D = NA,
                   Y_sim1 = NA,
                   y_sim1 = NA,
                   yhat_sim1 = NA,
                   c_sim1 = NA,
                   chat_sim1 = NA)
# Start Calculating Values
sim1 <- sim1 %>%
  mutate(L = L_0*exp(n*(year-2020)),
         A = A_0*exp(g*(year-2020)),
         m = x*((1-(sm/w))^epsilon)*Z[1],
         D = 1 - (1/(1+theta1*T[1]^theta2)))

sim1 <- sim1 %>%
  mutate(Y_sim1 = (1-D[1])*(Z[1]))

sim1 <- sim1 %>%
  mutate(y_sim1 = Y_sim1[1]/L_0,
         c_sim1 = (1-sk-sm)*y_sim1[1])

for (i in 2:nrow(sim1)){
  sim1$K[i] <- (1-delta)*sim1$K[i-1] + sk*sim1$Y_sim1[i-1]
  sim1$M[i] <- sim1$M[i-1] + sim1$m[i-1]
  sim1$Z[i] <- (sim1$K[i]^alpha)*(sim1$A[i]*sim1$L[i])^(1-alpha)
  sim1$m[i] <- x*(1-(sm/w))^epsilon*sim1$Z[i]
  sim1$T[i] <- 1 + ((sim1$M[i]^k_greek)/1000)
  sim1$D[i] <- 1 - (1/(1+theta1*sim1$T[i]^theta2))
  sim1$Y_sim1[i] <- (1-sim1$D[i])*(sim1$Z[i])
  sim1$y_sim1[i] <- sim1$Y_sim1[i]/sim1$L[i]
  sim1$c_sim1[i] <- (1-sk-sm)*sim1$y_sim1[i]
  sim1$yhat_sim1[i] <- log(sim1$y_sim1[i]/sim1$y_sim1[i-1])
  sim1$chat_sim1[i] <- log(sim1$c_sim1[i]/sim1$c_sim1[i-1])
}
```



# Simulation 2: Full abatment effort, `r s_m` = `r w`
# second sim Parameters 
```{r}
w <- 0.05
sm <- w
L_0 <- 7.9
K_0 <- 170
M_0 <- 500
A_0 <- 7
sk <- 0.25
delta <- 0.04
alpha <- 0.33
n <- 0.015
g <- 0.02
x <- 0.7
epsilon <- 0.5

k_greek <- 0.75
theta1 <- 0.05
theta2 <- 2
```
# Second simulation
```{r}
sim2 <- data.frame(year = 2020:2100,
                 L = L_0,
                   K = K_0,
                   A = A_0,
                   sm = sm,
                   Z = (K_0^alpha)*((A_0*L_0)^(1-alpha)),
                   m = NA,
                   M = M_0,
                   T = 1 + (M_0^k_greek/1000),
                   D = NA,
                   Y_sim2 = NA,
                   y_sim2 = NA,
                   yhat_sim2 = NA,
                   c_sim2 = NA,
                   chat_sim2 = NA)
# Start Calculating Values
sim2 <- sim2 %>%
  mutate(L = L_0*exp(n*(year-2020)),
         A = A_0*exp(g*(year-2020)),
         m = x*((1-(sm/w))^epsilon)*Z[1],
         D = 1 - (1/(1+theta1*T[1]^theta2)))

sim2 <- sim2 %>%
  mutate(Y_sim2 = (1-D[1])*(Z[1]))

sim2 <- sim2 %>%
  mutate(y_sim2 = Y_sim2[1]/L_0,
         c_sim2 = (1-sk-sm)*y_sim2[1])

for (i in 2:nrow(sim2)){
  sim2$K[i] <- (1-delta)*sim2$K[i-1] + sk*sim2$Y_sim2[i-1]
  sim2$M[i] <- sim2$M[i-1] + sim2$m[i-1]
  sim2$Z[i] <- (sim2$K[i]^alpha)*(sim2$A[i]*sim2$L[i])^(1-alpha)
  sim2$m[i] <- x*(1-(sm/w))^epsilon*sim2$Z[i]
  sim2$T[i] <- 1 + ((sim2$M[i]^k_greek)/1000)
  sim2$D[i] <- 1 - (1/(1+theta1*sim2$T[i]^theta2))
  sim2$Y_sim2[i] <- (1-sim2$D[i])*(sim2$Z[i])
  sim2$y_sim2[i] <- sim2$Y_sim2[i]/sim2$L[i]
  sim2$c_sim2[i] <- (1-sk-sm)*sim2$y_sim2[i]
  sim2$yhat_sim2[i] <- log(sim2$y_sim2[i]/sim2$y_sim2[i-1])
  sim2$chat_sim2[i] <- log(sim2$c_sim2[i]/sim2$c_sim2[i-1])
}
```
# Third simulation

# Third sim Parameters 
```{r}
w <- 0.05
sm <-  0.025           #sm(t=0), the new sm is added in the dataframe
L_0 <- 7.9
K_0 <- 170
M_0 <- 500
A_0 <- 7
sk <- 0.25
delta <- 0.04
alpha <- 0.33
n <- 0.015
g <- 0.02
x <- 0.7
epsilon <- 0.5
n_greek <- 0.02 #for sm
k_greek <- 0.75
theta1 <- 0.05
theta2 <- 2
```


# Third simulation
```{r}
sim3 <- data.frame(year = 2020:2100,
                 L = L_0,
                   K = K_0,
                   A = A_0,
                   sm = NA,
                   Z = (K_0^alpha)*((A_0*L_0)^(1-alpha)),
                   m = NA,
                   M = M_0,
                   T = 1 + (M_0^k_greek/1000),
                   D = NA,
                   Y_sim3 = NA,
                   y_sim3 = NA,
                   yhat_sim3 = NA,
                   c_sim3 = NA,
                   chat_sim3 = NA)
# Start Calculating Values
sim3 <- sim3 %>%
  mutate(L = L_0*exp(n*(year-2020)),
         A = A_0*exp(g*(year-2020)),
         sm = w/ (1+ exp(n_greek* (2020- year))),
         m = x*((1-(sm/w))^epsilon)*Z[1],
         D = 1 - (1/(1+theta1*T[1]^theta2)))

sim3 <- sim3 %>%
  mutate(Y_sim3 = (1-D[1])*(Z[1]))

sim3 <- sim3 %>%
  mutate(y_sim3 = Y_sim3[1]/L_0,
         c_sim3 = (1-sk-sm)*y_sim3[1])

for (i in 2:nrow(sim3)){
  sim3$K[i] <- (1-delta)*sim3$K[i-1] + sk*sim3$Y_sim3[i-1]
  sim3$M[i] <- sim3$M[i-1] + sim3$m[i-1]
  sim3$Z[i] <- (sim3$K[i]^alpha)*(sim3$A[i]*sim3$L[i])^(1-alpha)
  sim3$m[i] <- x*((1-(sim3$sm/w))^epsilon)*sim3$Z[i]
  sim3$T[i] <- 1 + ((sim3$M[i]^k_greek)/1000)
  sim3$D[i] <- 1 - (1/(1+theta1*sim3$T[i]^theta2))
  sim3$Y_sim3[i] <- (1-sim3$D[i])*(sim3$Z[i])
  sim3$y_sim3[i] <- sim3$Y_sim3[i]/sim3$L[i]
  sim3$c_sim3[i] <- (1-sk-sm)*sim3$y_sim3[i]
  sim3$yhat_sim3[i] <- log(sim3$y_sim3[i]/sim3$y_sim3[i-1])
  sim3$chat_sim3[i] <- log(sim3$c_sim3[i]/sim3$c_sim3[i-1])
}
```


# Evaluation analysis

step1: have a df of each simulation's consumption per capita
```{r}
cpc_df <- data.frame(year= 2020:2100, c1= sim1$c_sim1, c2= sim2$c_sim2, c3= sim3$c_sim3) 
```
Step2: cacluate the normal cpc difference between sim1&sim2, sim1&sim3 in the cpc_df, call them Current Value difference (CV diff)
```{r}
cpc_df <- cpc_df%>%
  mutate(
    CV_diff_c2_c1 = c2 -c1,
    CV_diff_c3_c1= c3 - c1,
  )
```
Step3: Caclulate the PV of  CV_diff_c1_c2, and  CV_diff_c1_c3
```{r}
#PV[x(t)]= exp(-r*(year-2020))*(CV_diff)
r <- 0.03
cpc_df <- cpc_df%>%
  mutate(
    PV_CV_diff_c2_c1= exp(-r*(2021-2020))*(c2[1]-c1[1]),
    PV_CV_diff_c3_c1= exp(-r*(2021-2020))*(c3[1]-c1[1]) )


for(i  in 1:nrow(cpc_df)){
  cpc_df$PV_CV_diff_c2_c1[i]<- exp(-r*(cpc_df$year[i]-2020))*(cpc_df$c2[i]-cpc_df$c1[i])
  cpc_df$PV_CV_diff_c3_c1[i]<- exp(-r*(cpc_df$year[i]-2020))*(cpc_df$c3[i]-cpc_df$c1[i])
}
```
Step4: Calculate the CPV for each PV_diff
```{r}
cpc_df <- cpc_df%>%
  mutate(
    CPV_PV_c2_c1 = cumsum(PV_CV_diff_c2_c1),
    CPV_PV_c3_c1 = cumsum(PV_CV_diff_c3_c1)
    
    
  )

```

# The problem
We differ with Sem's results for sim3&sim1, and that is because we have  different `r m`values. Either the results differ due to rounding issues, or due to something wrong with how we formulated it. 

This df below is a way to compare it with Sem's numbers, have Sem's excel next to you for compariosn
```{r}
c3_scrap <- sim3$c_sim3
c3_scrap_Df <- data.frame(c3_scrap)
c3_scrap_Df$c3_rounded <- round(c3_scrap_Df$c3_scrap, 2)
```








